---
title: "Ad-Hoc Requests"
output: 
  rmdformats::downcute:
    toc_depth: 3
---

This section summarizes and documents historic ad-hoc requests for geospatial, census, or marketing supporting data. 

# Ad-Hoc Requests

## DATA - 16

[DATA-16](https://solsticepower.atlassian.net/browse/DATA-16)

Request received on 1/8/2024. Summary was to pull Illinois data for Knox, Fulton, Peoria, Stark, Henry, Mercer, Warren and Winnebago Counties; in Minnesota pulling for Sherburne, Wright, Wabasha, Chisago, Meeker and any bordering counties. 

```{r setup, include=FALSE}

source('helper_functions.R')
load_packages()


```



```{r message=FALSE, warning=FALSE, include=FALSE}


il_final_df <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/ILSFA/il_final_table.csv")

mn_final_df <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/ILSFA/mn_final_table.csv")
```

### Illinois

```{r echo=FALSE, message=FALSE, warning=FALSE}

datatable(il_final_df, caption = "Illinois County Demographic Data Requested")%>%
    formatCurrency('Total Population',currency = "", interval = 3, mark = ",", digits = 0) %>%
    formatCurrency('Total HHs',currency = "", interval = 3, mark = ",", digits = 0) %>%
    formatCurrency('<80% AMI HHs',currency = "", interval = 3, mark = ",", digits = 0)


```

### Minnesota

```{r echo=FALSE, message=FALSE, warning=FALSE}

datatable(mn_final_df, caption = "Minnesota County Demographic Data Requested")%>%
    formatCurrency('Total Population',currency = "", interval = 3, mark = ",", digits = 0) %>%
     formatCurrency('Total HHs',currency = "", interval = 3, mark = ",", digits = 0) %>%
    formatCurrency('<150% AMI Pop',currency = "", interval = 3, mark = ",", digits = 0) %>%
    formatCurrency('<150% AMI HHs',currency = "", interval = 3, mark = ",", digits = 0) 


```











