---
title: "Delaware"
output: 
  rmdformats::downcute:
    toc_depth: 3
---


```{r setup, include=FALSE}

source('helper_functions.R')
load_packages()



```


# Overivew

Delaware's [Community Energy Facilities](https://depsc.delaware.gov/2022/02/01/community-energy-facilities-webpage/) program has an LMI carveout of 15%. LMI customers are defined via geoqualificaiton and income verification/self-attestation. Limited number and availability of utility data requires us to forego the utility overlays for Delaware. We will focus instead on estimating the total population considered for geo- and LMI qualification. 

# Geoqualification

Customers are defined as geoqualified if they reside in a census block "where the median household income is at or below 200% of the U.S. Federal Poverty Guidelines or 60% of the state median household income published by the United States Census Bureau, whichever is greater". 

Note, census blocks do not contain income information. This data is only available at the census block group or tract level. To approximate the population geoqualified in Delaware, we use a modified definition. Using 2021 5-Year ACS data, we define a geoqualified block group as one where at least 50% of the population is at or below 200% of the [Federal Poverty Line](https://data.census.gov/table?q=c17002&g=040XX00US10,10$1400000,10$1500000). 
    
There are 706 block groups in Delaware, a total of 77 are tagged as geoqualified using this method. The below map shows our tagged census block groups: 



```{r message=FALSE, warning=FALSE, include=FALSE}



get_blocks <- get_acs(
  geography="block group", 
  state="DE",
  variables=c("Total" = "C17002_001",
              "group1" = "C17002_002",
              "group2" = "C17002_003",
              "group3" = "C17002_004",
              "group4" = "C17002_005",
              "group5" = "C17002_006",
              "group6" = "C17002_007",
              "group7" = "C17002_008",
              "snap_hh" = "B19123_002",
              "snap_pop" = "B19058_002"),
  year=2021, 
  geometry=TRUE)  


####

get_blocks <- get_blocks %>%
  group_by(GEOID, geometry)  %>%
  summarize(total_pop = estimate[variable=="Total"],
            total_under_200FPL =
                          estimate[variable=="group1"] + 
                          estimate[variable=="group2"] +
                          estimate[variable=="group3"] +
                          estimate[variable=="group4"] +
                          estimate[variable=="group5"] +
                          estimate[variable=="group6"] ,
            snap_hh = estimate[variable=="snap_hh"],
            snap_pop = estimate[variable== "snap_pop"]) %>%
  mutate(pct_under_200FPL = total_under_200FPL / total_pop) %>%
  mutate(geo_tag = case_when(
    pct_under_200FPL >= 0.5 ~ 1,
    TRUE ~ 0
  ))

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
pal5 <- colorFactor(
  palette=c("lightgrey", "darkgreen"),
  domain=get_blocks$geo_tag
)


leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data=get_blocks,
              group="Block Groups",
             # stroke = TRUE,
              color = ~pal5(get_blocks$geo_tag),
              fillOpacity=0.75,
              popup=paste("Census Block Group: ", get_blocks$GEOID, "<br>",
                          "Percent under 200% FPL: ", get_blocks$pct_under_200FPL, "<br>",
                          "Number under 200% FPL: ", get_blocks$total_under_200FPL)) %>%
  addLegend(position="bottomright",
            pal=pal5,
            values=get_blocks$geo_tag,
            title="DE Geoqualified Census Block Groups") 

  
```


The total popoulation, population at or below 200% and SNAP households and population are summarized below. Approximately 25% of the total population in Delaware reside in census blocks geoqualified under the CEF program.  

```{r echo=FALSE, message=FALSE, warning=FALSE}

temp_table <- get_blocks %>% 
  st_drop_geometry() %>%
    mutate(Geoqualified = case_when(
    geo_tag == 1 ~ "Yes",
    TRUE ~ "No"
  )) %>%
  group_by(Geoqualified) %>%

  summarize(Population = sum(total_pop),
            `Population Under 200% FPL` = sum(total_under_200FPL),
            SNAP_HH = sum(snap_hh),
            SNAP_Pop = sum(snap_pop)) %>%
  adorn_totals("row")


datatable(temp_table, caption = "Delaware Geoqualified Population") %>%
    formatCurrency('Population',currency = "", interval = 3, digits = 0, mark = ",") %>%
    formatCurrency('Population Under 200% FPL',currency = "", interval = 3, digits = 0, mark = ",") %>%
    formatCurrency('SNAP_HH',currency = "", interval = 3, digits = 0, mark = ",") %>%
    formatCurrency('SNAP_Pop',currency = "", interval = 3, digits = 0, mark = ",")


```




# LMI Qualificaiton

There exists a number of options to verify customer income. Including:

    * (1) Proof of income of the account holder (e.g., pay stub or W2)
    * (2) proof of participation in a low income discount program including Medicaid, SSI, TANF, GA, WIC, LIHEAP, SNAP or food stamps
    * (3) proof that the Low-income Customer lives in a census block where the median household income is at or below 200% of the U.S. Federal Poverty Guidelines or 60% of the state median household income published by the United States Census Bureau, whichever is greater
    * (4) by living in a low-income master-metered building
    * (5) A written attestation by the Low-income Customer that their total household income is at or below 200% of the Federal Poverty Guidelines, or 60% of the state median household income published by the United States Census bureau, whichever is greater
    * NOTE For owners or operators of master-metered buildings, a written attestation that their tenants meet the income eligibility requirements for Low-income Customers
    
We use SNAP or food stamp recipient data as a proxy for low income discount programs. To keep consistency with the geoqualified tracts, we use block groups, hence Medicaid population is not available at this level. 


```{r echo=FALSE, message=FALSE, warning=FALSE}
pal5 <- colorNumeric(
  palette=c("Oranges"),
  domain=get_blocks$snap_pop
)

pal6 <- colorNumeric(
  palette=c("Greens"),
  domain=get_blocks$total_under_200FPL
)



leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data=get_blocks,
              group="SNAP",
             # stroke = TRUE,
              color = ~pal5(get_blocks$snap_pop),
              fillOpacity=0.75,
              popup=paste("Census Block Group: ", get_blocks$GEOID, "<br>",
                          "Percent under 200% FPL: ", get_blocks$pct_under_200FPL, "<br>",
                          "Number under 200% FPL: ", get_blocks$total_under_200FPL)) %>%
  addLegend(position="bottomright",
            pal=pal5,
            group="SNAP",
            values=get_blocks$snap_pop,
            title="DE SNAP Population")  %>%
  
  addPolygons(data=get_blocks,
              group="200% FPL",
             # stroke = TRUE,
              color = ~pal6(get_blocks$total_under_200FPL),
              fillOpacity=0.65,
              popup=paste("Census Block Group: ", get_blocks$GEOID, "<br>",
                          "Percent under 200% FPL: ", get_blocks$pct_under_200FPL, "<br>",
                          "Number under 200% FPL: ", get_blocks$total_under_200FPL)) %>%
  addLegend(position="bottomleft",
            pal=pal6,
            group= "200% FPL",
            values=get_blocks$total_under_200FPL,
            title="DE Total Under 200% FPL")  %>%
  
    addLayersControl(
    overlayGroups=c( "SNAP", "200% FPL"),  # Update the order for consistency
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup("SNAP")

  

  
```




# LIFT Solar


No projects in LIFT solar database in Delaware. 




