---
title: "Home"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::downcute:
    toc_depth: 3
---


```{r setup, include=FALSE}

source('helper_functions.R')
load_packages()
library(readxl)




```




# Overview

This document summarizes the total addressable markets within each state Solstice operates or plans to explore. 

Each state defines eligible population for solar farms through unique programmatic details, listed in detail on the [Policy Matrix](https://coda.io/d/Solstice-Policy_dBUL-zOSfnb/Program-Matrix_suVJB#Program-Matrix_tuLIr/r11&view=full) on Coda. This work looks to define how many people are eligible for each state program via income, geographic, or other unique definitions. Many programs have LMI carveouts; this document looks to define how many people that equates to. When available and applicable, utiliy zones are overlaid with these populations to estimate the potential population Solstice would be competing to sign up. 

Utility coverage is provided by either the [HIFLD](https://hifld-geoplatform.opendata.arcgis.com/datasets/f4cd55044b924fed9bc8b64022966097/explore?location=41.065293%2C-80.583661%2C6.90) national level electric retail service territories or, preferably, state sources that are more granular. Note HIFLD often provides overlapping areas (ex: a municipal boundary and an utility provider) so data should primarily be sourced from state sources or be validated from HIFLD. 

# Parcel Coverage

Using parcel data, we can tag individual plots considered eligible by various state programs. When data is available, further analyses can provide breakdowns between residential and commercial buildings, allowing marketing to zero in on potential sites for a variety of applications such as corporation identification. Note, this process is only available in states where sufficient and open sourced parcel data is made available. Examples include [Illinois](https://jake-ford.github.io/lmi_by_utility_v2/Illinois.html). 

# Solar Saturation

We use the Groundswell's [LIFT](https://lift.groundswell.org/solar-projects/) to analyze existing CS market sizes and compare these to total capacity in the current state to arrive at existing capacity. State specific resources may supplement LIFT. 



# Total Eligible Population Comparison

```{r echo=FALSE, message=FALSE, warning=FALSE}

ca_geo <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/ca_geo_total.csv") %>%
 # mutate(State = "California") %>%
  select( Population, Percentage)

il_geo_sfa <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/il_geo_sfa_total.csv")%>%
 # mutate(State = "Illinois SFA")  %>%
  select( Population, Percentage)


il_geo_engie <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/il_geo_engie_total.csv")%>%
 # mutate(State = "Illinois Engie")  %>%
  select( Population, Percentage)


nj_geo <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/nj_geo_total.csv")%>%
#  mutate(State = "New Jersey")  %>%
  select( Population, Percentage)


ny_geo <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/ny_geo_total.csv")%>%
 # mutate(State = "New York")  %>%
  select( Population, Percentage)


ma_geo <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/ma_geo_total.csv") %>%
  mutate(State = "Massachusetts",
         Percentage = `Geo-Qual Population`/`State Total Pop`) %>%
  select(State, `Geo-Qual Population`, Percentage, `80% AMI`, `100% AMI`, SNAP = `SNAP HHs`, `Total Pop` = `State Total Pop`, -Utility)



temp_geo <- rbind(ca_geo, il_geo_sfa, il_geo_engie, nj_geo, ny_geo)



ca_lmi <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/ca_lmi_total.csv") %>%
  mutate(State = "California") %>%
  select(-Utility)

cal_total <- cbind(ca_geo, ca_lmi)

il_lmi <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/il_lmi_total.csv")%>%
  mutate(State = "Illinois") %>%
  select(-Utility)

il_total_sfa <- cbind(il_geo_sfa, il_lmi)
il_total_sfa$State <- "Illinois SFA"

il_total_engie <- cbind(il_geo_engie, il_lmi)
il_total_engie$State <- "Illinois Engie"


nj_lmi <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/nj_lmi_total.csv")%>%
  mutate(State = "New Jersey") %>%
  select(-Utility)

nj_total <- cbind(nj_geo, nj_lmi)

nj_total$`80% AMI` <- 1144376
nj_total$`100% AMI` <- 1639804
nj_total$SNAP <- 280379
nj_total$`Total Pop` <- nj_total$`State Total Pop`

nj_total <- nj_total %>%
  select(Population, Percentage, `80% AMI`, `100% AMI`,
         SNAP, `Total Pop`, State)


ny_lmi <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/ny_lmi_total.csv")%>%
  mutate(State = "New York") %>%
  select(-Utility)

ny_total <- cbind(ny_geo, ny_lmi)


nm_lmi <- read_csv("/Users/jakeford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/Geospatial Locational Analysis/market_analysis/index_data/nm_lmi_total.csv") %>%
  mutate(State = "New Mexico") %>%
  rename(`Total Pop` = `Total Population`,
         `SNAP` = `SNAP HH`) %>%
  select(-Name, -Medicaid)

nm_lmi$Population <- NA
nm_lmi$Percentage <- NA

cols <- names(nm_lmi)
nm_lmi <- nm_lmi[, c("Population", "Percentage", cols[!(cols %in% c("Population", "Percentage"))])]


temp_geo <- rbind(cal_total, il_total_sfa,
                  il_total_engie, nj_total, ny_total, nm_lmi)

temp_geo <- temp_geo[, c("State", setdiff(names(temp_geo), "State"))]

temp_geo <- temp_geo %>% 
  rename(`Geo-Qual Population` = Population)

temp_geo <- rbind(temp_geo, ma_geo)


datatable(temp_geo, caption = "Solstice State Markets Geo- and LMI Populations") %>%
    formatCurrency('80% AMI',currency = "", interval = 3, digits = 0, mark = ",") %>%
    formatCurrency('Geo-Qual Population',currency = "", interval = 3, digits = 0, mark = ",") %>%
    formatCurrency('100% AMI',currency = "", interval = 3, digits = 0, mark = ",") %>%   formatCurrency('SNAP',currency = "", interval = 3, digits = 0, mark = ",") %>%
  formatCurrency('Total Pop',currency = "", interval = 3, digits = 0, mark = ",") %>%
  formatPercentage('Percentage', digits=2)


```








# IRA Eligible Population by State

This analysis provides an initial state level estimate of the population eligible for the IRA bonus tax credit as described on the Solstice [website](https://solstice.us/solstice-blog/inflation-reduction-act-investment-tax-credit/). As a reminder, eligible populations must reside in a Persistent Poverty County or in a Climate & Economic Justice Screening Tool: census tracts designated as disadvantaged based on 90th percentile energy burden or PM2.5 exposure + 65th percentile low-income. 

In addition to the IRA bonus tax credit, it is useful to consider the New Markets Tax Credit (NMTC) program for a comprehensive understanding of economic incentives in low-income and distressed communities. The NMTC program, managed by the Community Development Financial Institutions (CDFI) Fund, targets investment in communities that meet specific low-income criteria, primarily based on poverty rates or median family income. Unlike the IRA, which focuses on specific environmental and economic factors such as energy burden or PM2.5 exposure, the NMTC program centers more broadly on economic revitalization. Eligibility under the NMTC program is determined by census tracts where the poverty rate is at least 20% or where median family income does not exceed 80% of the area median income (AMI). This approach ensures that investments are directed towards areas in need of economic growth and infrastructure development, complementing the objectives of the IRA. Understanding the overlap and distinct features of both the IRA and NMTC can provide a more holistic view of the socio-economic landscape and opportunities for investment in underserved communities.

The below table shows the following:


* **Population** - Total 2022 ACS 5 Year Population.
* **NMTC_2015** - Proportion of population covered by the 2023 Census tract boundary that satisfies the definition of a low-income community for the purposes of Category 1 applications via the New Market Tax Credit (NMTC) map created using 2011-2015 ACS data. Facilities in NMTC_2015 regions meet the geographic eligibility criteria for Category 1 only if applications are submitted *before* August 31st, 2024.
* **NMTC_2020** - Proportion of the population covered by the 2023 Census tract boundary that satisfies the definition of a low-income community for the purposes of Category 1 applications via the New Market Tax Credit (NMTC) map created using 2016-2020 ACS data. 
* **CEJST** - Proportion of population covered that meets the Climate and Economic Justice Screening Tool's (CEJST) threshold for disadvantage in the 'Energy' category of Burden using data released by the CEJST Program in November 2022.
* **PPC** - Persistent Poverty County population as defined by the EDA. 


```{r message=FALSE, warning=FALSE, include=FALSE}


# Good map here of [census tract](https://experience.arcgis.com/experience/12227d891a4d471497ac13f60fffd822/page/Page/?data_id=dataSource_3-18add75d2c8-layer-6%3A34787) and full data [source](https://data.nrel.gov/submissions/222#:~:text=Counties%20are%20displayed%20that%20meet,plus%203%20evaluation%20time%20periods). 

tract_data <- read_excel("~/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/IRA_Analysis/qualified_tracts_overlay.xlsx",
                         sheet = "Tract eligibility percentages") %>%
  mutate(GEOID = GEOID_2022)

ppc_data <- read_excel("~/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/IRA_Analysis/EDA_FY23_PPCs.xlsx",
                         sheet = "Underlying_Data") %>%
  mutate(ppc_tag = ...7)




```


```{r message=FALSE, warning=FALSE, include=FALSE}



# Make sure your API key is set (replace 'your_api_key' with your actual key)
#census_api_key("ad86ad3edc7d52897ffc0f251537babfddeac870", install = TRUE, overwrite=TRUE)

# Get the vector of state abbreviations
states <- state.abb

# Initialize an empty data frame to store results
total_population <- data.frame()

# Loop through each state
for (state in states) {
  # Attempt to request tract-level data for the current state
  state_data <- tryCatch({
    get_acs(geography = "tract", 
            variables = "B01001_001", 
            geometry = FALSE, 
            year = 2022, 
            state = state)
  }, error = function(e) {
    # Print the error message
    message("Error for state: ", state, "; Error: ", e$message)
    # Return NULL in case of an error
    return(NULL)
  })
  
  # If data is retrieved, bind it to the total_population data frame
  if (!is.null(state_data)) {
    total_population <- rbind(total_population, state_data)
  }
}

# Check the result
if(nrow(total_population) == 0) {
  message("No data was retrieved.")
} else {
  message("Data retrieved successfully.")
}

```


```{r echo=FALSE, message=FALSE, warning=FALSE}


# Check if the columns exist and are unique
if(!"GEOID" %in% names(tract_data)) stop("Column 'GEOID_2022' not found in tract_data")
if(!"GEOID" %in% names(total_population)) stop("Column 'GEOID' not found in total_population")
if(anyDuplicated(tract_data$GEOID_2022)) stop("Duplicate values found in 'GEOID_2022' of tract_data")
if(anyDuplicated(total_population$GEOID)) stop("Duplicate values found in 'GEOID' of total_population")

# Perform the merge
merged_df <- merge(tract_data, total_population, by="GEOID")
#merged_df <- merge(tract_data, total_population, by.x = "GEOID_2022", by.y = "GEOID")

# Rest of your code

#merged_df <- merge(tract_data, total_population, by.x="GEOID_2022", by.y= "GEOID") 

merged_df$StateName <- tolower(merged_df$StateName)

merged_df$StateAbbreviation <- state.abb[match(merged_df$StateName, tolower(state.name))]



# Create a vector of names from ppc_data where ppc_tag is "Yes"
names_with_yes_tag <- ppc_data$Name[ppc_data$ppc_tag == "Yes"]

merged_df <- merged_df %>%
  mutate(Name = paste(CountyName, "County,", StateAbbreviation)) %>%
  mutate(ppc_tag = as.integer(Name %in% names_with_yes_tag)) 





```




```{r echo=FALSE, message=FALSE, warning=FALSE}


flipped_df <- merged_df %>%
  group_by(State = StateName) %>%
  summarize(Population = sum(estimate, na.rm=T),
            NMTC_2015 = round(sum(NTMC2015_pct * estimate), 0 ),
            NMTC_2020 = round(sum(NTMC2020_pct * estimate), 0 ),
            CEJST = round(sum(CEJST2022_pct * estimate), 0 ),
            PPC = round(sum(ppc_tag * estimate), 0 )
            ) %>%
 # mutate(NMTC_2022 = NTMC2020_pct * Population) %>%
  arrange(desc(Population)) %>%
  adorn_totals("row")


datatable(flipped_df, caption = "State IRA Eligible Populations") %>%
    formatCurrency('Population',currency = "", interval = 3, mark = ",", digits = 0) %>%
    formatCurrency('NMTC_2015', currency = "", interval = 3, mark = ",", digits=0) %>%
    formatCurrency('NMTC_2020', currency = "", interval = 3, mark = ",", digits=0) %>%
    formatCurrency('CEJST', currency = "", interval = 3, mark = ",", digits=0) %>%
    formatCurrency('PPC', currency = "", interval = 3, mark = ",", digits=0)




```





















